---
title: "OJ"
output: html_document
---

```{r setup, include = FALSE}
# import libraries
library(ggplot2)
library(tidyverse)
library(dplyr)
```

```{r load-data}
# read in data
oj <- read.csv('oj.csv')
```

## Visualizing price  


Make a box plot of price.
```{r box-plot-price}
oj %>% 
  ggplot(aes(x = "", y = price)) +
  geom_boxplot()
```


Make a box plot of log price.
```{r box-plot-log-price}
oj <- oj %>%
  mutate(log_price = log(price)) 

oj %>% 
  ggplot(aes(x = "", y = log_price)) +
  geom_boxplot()
```


Make a box plot of price, but separate out each brand.
```{r box-plot-price-brands}
oj %>% 
  ggplot(aes(factor(brand), price)) +
  geom_boxplot(aes(fill = factor(brand)))
```


Do the same for log price. 
```{r box-plot-log-price-brands}
oj %>% 
  mutate(log_price = log(price)) %>%
  ggplot(aes(factor(brand), log_price)) +
  geom_boxplot(aes(fill = factor(brand)))
```


#### What do these graphs tell you about the variation in price? Why do the log plots look different? Do you find them more/less informative?

The log plots are more informative in terms of visualizing variance. We can more easily see how the outliers are spread out.


## Visualizing the quantity/price relationship

Plot logmove (log quantity) vs. log(price) for each brand.  
```{r quantity-price}
oj %>%
  mutate(log_price = log(price)) %>%
  ggplot(aes(x = logmove, y = log_price)) +
  geom_point(aes(color = brand))
```

#### What do insights can you derive that were not apparent before?

In contrast to the box plot, here we can see how price affects the quantity
of orange juice sold for each brand. For all 3 brands, with an increase in
price, fewer products are sold. 


## Estimating the relationship

Do a regression of log quantity on log price.
```{r estimating-relationship-a}

regression <- oj %>%
  lm(formula = logmove ~ log_price)

summary(regression)
```

#### How well does the model fit? What is the elasticity, does it make sense?

R-squared is small => doesn't explain variance well, so the model is not a great fit.




Now add in an intercept term for each brand (add brand to the regression), how do the results change? 

``` {r estimating-relationship-b}
regression_with_intercept <- oj %>%
  lm(formula = logmove ~ log_price + brand)

summary(regression_with_intercept)
```


Now figure out a way to allow the elasticities to differ by brand. Search 
"interaction terms" and "dummy variables" if you don't remember this from 
econometrics. Note the estimate coefficients will "offset" the base estimates.
What is the insights we get from this regression? What is the elasticity for each 
firm? Do the elasticities make sense?

```{r estimating-relationship-c}
regression_with_interaction <- oj %>%
  lm(formula = logmove ~ log_price * brand)

summary(regression_with_interaction)


# plot(logmove ~ log_price, data = oj)
# abline(regression, col = "blue")
# abline(regression_with_intercept, col = "darkgreen")
# abline(regression_with_interaction, col = "red")

```


Hold out 20% of your sample randomly.  Estimate the model on the remaining 80%.
Use the predict command to see how well the model fits on the rest of the 
data (e.g., y_hat <- predict(my_model, newx = test_matrix))
```{r predict}
  
idx <- as.integer((nrow(oj)*0.8))

set.seed(42)
train_idx <- sample(seq_len(nrow(oj)), idx)

train <- data.frame(oj[train_idx,])
test <- data.frame(oj[-train_idx,])

model <- lm(formula = logmove ~ log_price + brand + log_price*brand, data = oj)

y_hat <- predict(model, newx = test)
summary(y_hat)


```

## Impact of "featuring in store"

Which brand is featured the most? 
```{r feature-a}
oj %>%
  group_by(store, brand, feat) %>% summarize(count= n()) %>% filter(feat == 1) %>%
  ggplot(aes(x = store, y = count)) +
  geom_point(aes(color = factor(brand)), position = "jitter") +
  xlab("Store ID") +
  ylab("Number of Times Featured")

oj %>%
  ggplot(aes(feat, brand)) +
  geom_point(position = "jitter", aes(color = brand))

```

What is the average price and featured rate of each brand? 
```{r feature-a-i}

oj %>% 
  group_by(brand) %>% 
  summarize(feat = mean(feat), price = mean(price))

oj %>%
  group_by(brand, feat) %>%
  summarize(mean_price = mean(price), sd_price = sd(price), obs = length(price))

```



How should incorporate the feature variable into our regression? 
Start with an additive formulation (e.g. feature impacts sales, but not through price).
```{r feature-b}

regression_with_feature <- oj %>%
  lm(formula = logmove ~ log_price * brand + feat)

summary(regression_with_feature)

```

Now run a model where features can impact sales and price sensitivity.
```{r feature-c}

regression_with_feature_price <- oj %>%
  lm(formula = logmove ~ log_price * brand + feat * log_price)

summary(regression_with_feature_price)

```


Now add what you think are the most relevant sociodemographic controls and 
produce the regression results from that regression as well. 
```{r feature-d}
 
regression_with_socio <- oj %>%
  lm(formula = logmove ~ log_price * brand + feat * log_price + AGE60)

summary(regression_with_socio)
### R-squared = 0.5401

```
