---
title: "Citibike Maps"
author: "Sasha Paulovich"
date: "July 5, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*with help from https://rpubs.com/jhofman/nycmaps*

```{r libraries, message=FALSE, warning=FALSE}
library(tidyverse)
library(tigris)
library(dplyr)
library(leaflet)
library(sp)
library(ggmap)
library(maptools)
library(broom)
library(httr)
library(rgdal)
library(data.table)
library(lubridate)
```

#### Create a data frame that has the unique name, latitude, and longitude for each Citibike station that was present in the system in July 2014
```{r new-df}
july2014 <- read_csv('../week1/201407-citibike-tripdata.csv')
lat_long <- july2014 %>% select('start station name', 'start station latitude', 'start station longitude') %>% distinct()
setnames(lat_long, old=c("start station name", "start station latitude", "start station longitude"), new=c("station name", "station latitude", "station longitude"))
```

#### Make a map showing the location of each Citibike station using ggmap
```{r ggmap-lat-long}
nyc_map <- get_map(location = c(lon = -74.00, lat = 40.71), maptype = "terrain", zoom = 11)
ggmap(nyc_map) +
  geom_point(data=lat_long, aes(x=`station longitude`, y=`station latitude`), color="blue", size = 1, stroke = 0, shape = 16)
```

#### Do the same using leaflet, adding a popup that shows the name of the station when it's clicked on
```{r leaflet-popup}
bike_icon <- makeIcon('bike-icon.png', iconWidth = 40, iconHeight = 40)
leaflet() %>%
  addTiles() %>%
  setView(-74.00, 40.71, zoom = 12) %>%
  addProviderTiles("CartoDB.Positron") %>%
  addMarkers(lat_long$`station longitude`, lat_long$`station latitude`, 
             icon = bike_icon,
             popup = lat_long$`station name`)
```

#### Then do a spatial join to combine this data frame with the Pediacities NYC neighborhood shapefile data
```{r pediacities-data}
r <- GET('http://data.beta.nyc//dataset/0ff93d2d-90ba-457c-9f7e-39e47bf2ac5f/resource/35dd04fb-81b3-479b-a074-a27a37888ce7/download/d085e2f8d0b54d4590b1e7d1f35594c1pediacitiesnycneighborhoods.geojson')
nyc_neighborhoods <- readOGR(content(r,'text'), 'OGRGeoJSON', verbose = F)

# nyc_neighborhoods_df <- tidy(nyc_neighborhoods)
# ggmap(nyc_map) + 
  # geom_polygon(data=nyc_neighborhoods_df, aes(x=long, y=lat, group=group), color="blue", fill=NA)

lat_long_spdf <- lat_long
coordinates(lat_long_spdf) <- c("station longitude", "station latitude")
proj4string(lat_long_spdf) <- proj4string(nyc_neighborhoods)
matches <- over(lat_long_spdf, nyc_neighborhoods)

lat_long <- cbind(lat_long, matches)
```

#### Make a map showing the number of unique Citibike stations in each neighborhood
##### First do this using ggmap where the fill color encodes the number of stations
```{r ggmap-neighborhood-stations, warning=FALSE}
neighborhood_stations <- lat_long %>%
  group_by(neighborhood) %>%
  summarize(num_stations = n())

map_data <- geo_join(nyc_neighborhoods, neighborhood_stations, "neighborhood", "neighborhood")

plot_data <- tidy(nyc_neighborhoods, region="neighborhood") %>%
  left_join(., neighborhood_stations, by=c("id"="neighborhood")) %>%
  filter(!is.na(num_stations))

manhattan_map <- get_map(location = c(lon = -74.00, lat = 40.77), maptype = "terrain", zoom = 12)

ggmap(manhattan_map) +
  geom_polygon(data=plot_data, aes(x=long, y=lat, group=group, fill=num_stations), color="white", alpha=0.8)
```

##### Then do the same using leaflet, adding a popup that shows the number of stations in a neighborhood when its shape is clicked on
```{r leaflet-num-stations}
pal <- colorNumeric(palette = "PuRd",
                    domain = range(map_data@data$num_stations, na.rm=T))

leaflet(map_data) %>%
  addTiles() %>% 
  addPolygons(fillColor = ~pal(num_stations), fillOpacity = 0.4, 
              popup = paste("<b>", map_data@data$neighborhood, "</b>", "<br/>", 
                            "Number of Stations: ", map_data@data$num_stations)) %>% 
  addProviderTiles("CartoDB.Positron") %>%
  setView(-74.00, 40.71, zoom = 12)

```

#### Now create a new data frame that has the total number of trips that depart from each station at each hour of the day on July 14th
```{r july-14-df}
july14 <- july2014 %>% 
  mutate(day = day(as.Date(starttime)), hour = hour(starttime)) %>%
  filter(day == 14)
```


#### Do a spatial join to combine this data frame with the Pediacities NYC neighborhood shapefile data
```{r july14-neighborhood-df}
july14_spdf <- july14
coordinates(july14_spdf) <- c("start station longitude", "start station latitude")
proj4string(july14_spdf) <- proj4string(nyc_neighborhoods)
matches <- over(july14_spdf, nyc_neighborhoods)

july14 <- cbind(july14, matches)

july14_trips <- july14 %>%
  group_by(`start station name`, `start station longitude`, `start station latitude`, neighborhood, hour) %>%
  summarize(num_trips = n())

```


#### Make a ggmap plot showing the number of trips that leave from each neighborhood at 9am, 1pm, 5pm, and 10pm, faceted by hour, where each facet contains a map where the fill color encodes the number of departing trips in each neighborhood
```{r july14-ggmap, warning=FALSE}
map_data <- geo_join(nyc_neighborhoods, july14_trips, "neighborhood", "neighborhood")

plot_data <- tidy(nyc_neighborhoods, region="neighborhood") %>%
  left_join(., july14_trips, by=c("id"="neighborhood")) %>%
  filter(hour %in% c(9, 13, 17, 22), !is.na(num_trips)) %>%
  group_by(hour) %>%
  mutate(percentage = num_trips/sum(num_trips))

# plot_data$new_hour <- plot_data$hour %% 12
# plot_data$new_hour <- paste(plot_data$new_hour, ":00")

manhattan_map <- get_map(location = c(lon = -74.00, lat = 40.77), maptype = "terrain", zoom = 12)

ggmap(manhattan_map) +
  geom_polygon(data=plot_data, aes(x=long, y=lat, group=group, fill=num_trips), color="white", alpha=0.9) +
  scale_fill_gradient(low = "#375e9b", high = "red") +
  facet_wrap(~hour, nrow=1) + 
  labs(fill='Number of Trips') 
```
