---
title: "Trips from Penn Station to Grand Central (2014)"
author: "Sasha Paulovich"
date: "July 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries, message=FALSE}
library(tidyverse)
library(lubridate)
library(dplyr)
```

#### Load 2014 Citibike Data
```{r load-data}
load('../week1/trips.RData')
```

#### Create a function that computes historical trip times between any two stations:
* Take the trips dataframe and two station names as inputs  
* Return a 168-by-6 dataframe with summary statistics of trip times for each hour of the week (e.g., Monday 9am, Monday 10am, etc.), where the summary statistics include:  
+ Average number of trips in that hour  
+ Average and median trip times for that hour  
+ Standard deviation in trip time for that hour  
+ Upper and lower quartiles of trip time for that hour  

```{r}
summarize_stats <- function(station1, station2) {
  
  # filter for relevant trips
  df <- trips %>% 
    filter(start_station_name == station1, end_station_name == station2) %>%
    mutate(week = week(starttime), DOW = weekdays(starttime), hour = hour(starttime)) %>%
    select(starttime, week, DOW, hour, start_station_name, end_station_name, tripduration)
  
  # compute trip duration stats
  df_tripduration <- df %>% 
    group_by(DOW, hour) %>%
    summarize(avg_time = mean(tripduration),
              lower_quartile = quantile(tripduration, 0.25),
              median_time = median(tripduration),
              upper_quartile = quantile(tripduration, 0.75),
              sd = sd(tripduration)) 

  # compute average number of trips
  df_avgtrips <- df %>%
    group_by(week, DOW, hour) %>%
    summarize(count = n()) %>%
    ungroup() %>%
    group_by(DOW, hour) %>%
    summarize(mean_trips = mean(count))

  # combine the dfs
  df_stats <- full_join(df_avgtrips, df_tripduration)
  
  return(df_stats)
}


```


#### Use this function on trips between Penn Station and Grand Central (you can use the most popular station at each location)
```{r}
penn_station <- "8 Ave & W 31 St"
grand_central <- "E 43 St & Vanderbilt Ave"

ps_gc_stats <- summarize_stats(penn_station, grand_central)
```

#### Make a plot of the results, where each facet is a day of the week, the x axis shows hour of the day, and the y axis shows average trip time, with transparent ribbons to show the standard deviation in trip time around the mean
```{r}
ps_gc_stats %>%
  ggplot(aes(x = hour, y = avg_time)) +
  geom_line() +
  geom_ribbon(aes(ymin = avg_time - sd, ymax = avg_time + sd), alpha = 0.25) +
  facet_wrap(~DOW)
```
