---
title: 'Card & Krueger: Difference in Diffferences'
date: "June 28, 2018"
output: html_document
---

```{r setup}
library(tidyverse)

data <- read.csv('card_krueger_data.csv')

```


### Which outcome metrics could be impacted by the minimum wage change?

**EMPFT**: need to cut down on staff -- full time workers are expensive!   
**EMPPT**: perhaps more people will be applying, but fewer hires  
**WAGE_ST**: trivial  
**FIRSTINC**: starting wage will already be hire, so a raise will be less urgent  
**MEALS**: less of a discount to compensate for the higher wages?  
**PSODA**: trick customers into spending more money!!  
**PENTREE**: raise prices to compensate for higher wages


### Compute the 'diffs'
```{r diffs}

# rename for cleaner pattern matching
colnames(data)[colnames(data)=="FIRSTIN2"] <- "FIRSTINC2"

# get complementary indices and compute differences
for(col in colnames(data)[11:27]){
  col2 <- grep(paste("^", col, "[2]", sep=""), colnames(data))
  if(!is.na(col2[1])){
    data[[paste(col, "_DIFF", sep="")]] <- as.vector((data[col2] - data[col])[,1])
  }
}

# run t-tests on important features for NJ
nj <- data %>% filter(STATE == 1)
t.test(nj$WAGE_ST_DIFF)
t.test(nj$EMPFT_DIFF)
t.test(nj$EMPPT_DIFF)
t.test(nj$PENTREE)

# show just the means (Intercepts)
# colMeans(pa[47:61], na.rm=TRUE)

# run t-tests on important features for PA
pa <- data %>% filter(STATE == 0)
t.test(pa$WAGE_ST_DIFF)
t.test(pa$EMPFT_DIFF)
t.test(pa$EMPPT_DIFF)
t.test(pa$PENTREE)

# show just the means (Intercepts)
# colMeans(pa[47:61], na.rm=TRUE)


```


### Compute the "diff-in-diff" via the regression method
```{r diff-in-diff}
wage_st_reg <- lm(WAGE_ST_DIFF ~ factor(STATE), data = data)
summary(wage_st_reg)

empft_reg <- lm(EMPFT_DIFF ~ factor(STATE), data = data)
summary(empft_reg)

emppt_reg <- lm(EMPPT_DIFF ~ factor(STATE), data = data)
summary(emppt_reg)

pentree_reg <- lm(PENTREE_DIFF ~ factor(STATE), data = data)
summary(pentree_reg)
```
### Did this analysis provide definitive evidence on the impact of the minimum wage on employment?
Yes, as we observe an increase in full time employment in NJ after the increase in minimum wage.

### Using an Independent Variable, GAP
```{r gap}
data <- data %>%
  mutate(GAP = (5.05 - WAGE_ST) / WAGE_ST)

gap_wage_reg <- lm(WAGE_ST_DIFF ~ as.factor(STATE)*GAP, data = data)
summary(gap_wage_reg)

gap_empft_reg <- lm(EMPFT_DIFF ~ as.factor(STATE)*GAP, data = data)
summary(gap_wage_reg)

gap_emppt_reg <- lm(EMPPT_DIFF ~ as.factor(STATE)*GAP, data = data)
summary(gap_wage_reg)
```
